\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-11-19 chapters/bttda.tex
Quantify qualitative statements in bttda chapter

\cref{tab:mi-score} lists the cross-validated classification scores for
the evaluated motor imagery datasets.
Note that, in line with~\textcite{Chevallier2024}, accuracy is reported for the
multi-class classification problem in BNCI2014-001, while ROC-AUC was reported for the
binary classification problem in BNCI2014-004.
\begin{table}[t]
  \centering
	\footnotesize
	\input{chapters/bttda/score_mi.tex}
  \caption[Within-session classification score for 2 motor imagery datasets.]{Classification score for
		cross-validated within-session evaluation for HODA+LDA and our proposed decoders
		PARAFACDA+LDA and BTTDA+LDA evaluated on 2 motor imagery datasets.
		Scores for other decoders were taken from \textcite{Chevallier2024}. Accuracy
    is listed for BNCI2014-001, and area under the receiver-operator
    characteristic curve for BNCI2014-004. BTTDA outperforms HODA and PARAFACDA
    for BNCI2014-001 but does not reach performance comparable to the current
    state-of-the-art. For BNCI2014-004, PARAFACDA and BTTDA perform
    approximately on par with HODA and the current state-of-the art.}
	\label{tab:mi-score}
\end{table}
For the BNCI2014-001 dataset, HODA+LDA and our proposed decoders do not reach
satisfactory performance compared to the comparison methods, yet both
PARAFACDA+LDA and BTTDA+LDA improve upon HODA+LDA, and for BTTDA+LDA this
difference is significant ($p=0.0269$).
For the BNCI2014-004 dataset, the performance gap with comparison methods is
smaller, but no significant differences were found between HODA+LDA,
PARAFACDA+LDA, and BTTDA+LDA\todo{qualitative}.

@@

Nevertheless, when considering the relative improvement over the original
HODA model, our results show that BTTDA+LDA performs consistently on par or
higher than HODA+LDA\todo{qualitative}.
The classification score for BTTDA+LDA was also slightly
higher\todo{qualitative} than that of
restricted PARAFACDA+LDA for the ERP datasets and one of the MI datasets, but these
results were not significant, and further studies with more datasets and
subjects should show whether this holds.
\Cref{fig:blocks} shows that there is added value in finding extra blocks
over the first HODA block.
While no proof is given, we notice that NMSE monotonically decreases, indicating that
since the training score keeps increasing, this suggests that
eventually all the variation in the signal will be explained by the model
while still extracting features that are maximally discriminant.
Eventually, the number of blocks will reach a point of diminishing validation
score returns, when adding extra features to the decision classifier increases
its risk of overfitting instead of adding extra useful discriminatory
information.
Together with the improved classification scores presented, these results
point to the potential of our more flexible model block-term or
its special PARAFAC-structured case over a Tucker-structured model given proper block and
rank selection.
Since the optimal ranks for HODA+LDA were also determined through
cross-validation, BTTDA+LDA can improve over the first HODA block.
If validation shows that this is not possible, such as is the case for
BNCI2014-004, the BTTDA model is truncated to the first HODA block and little
performance is lost.

@@

Via an accompanying heuristic model selection procedure, \ac{bci} decoders using
\ac{bttda} feature extraction can significantly outperform decoders based on
\ac{hoda} and reach state-of-the-art decoding performance on event-related potential
problems (second-order tensors) and scores on par with or higher than \ac{hoda} in motor
imagery problems (third-order tensors)\todo{qualitative}.
Moving from the rigid Tucker tensor structure of \ac{hoda} to the more flexible
block-term structure shifts the problem from finding optimally constrained
multilinear projections to model and feature selection.

>>>>

\cref{tab:mi-score} lists the cross-validated classification scores for
the evaluated motor imagery datasets.
Note that, in line with~\textcite{Chevallier2024}, accuracy is reported for the
multi-class classification problem in BNCI2014-001, while ROC-AUC was reported for the
binary classification problem in BNCI2014-004.
\begin{table}[t]
  \centering
	\footnotesize
	\input{chapters/bttda/score_mi.tex}
  \caption[Within-session classification score for 2 motor imagery datasets.]{Classification score for
		cross-validated within-session evaluation for HODA+LDA and our proposed decoders
		PARAFACDA+LDA and BTTDA+LDA evaluated on 2 motor imagery datasets.
		Scores for other decoders were taken from \textcite{Chevallier2024}. Accuracy
    is listed for BNCI2014-001, and area under the receiver-operator
    characteristic curve for BNCI2014-004. BTTDA outperforms HODA and PARAFACDA
    for BNCI2014-001 but does not reach performance comparable to the current
    state-of-the-art. For BNCI2014-004, PARAFACDA and BTTDA perform
    approximately on par with HODA and the current state-of-the art.}
	\label{tab:mi-score}
\end{table}
For the BNCI2014-001 dataset, HODA+LDA and our proposed decoders do not reach
satisfactory performance compared to the comparison methods, yet both
PARAFACDA+LDA and BTTDA+LDA improve upon HODA+LDA, and for BTTDA+LDA this
difference is significant ($p=0.0269$).
For the BNCI2014-004 dataset, the performance gap with comparison methods is
smaller, but no significant differences were found between HODA+LDA,
PARAFACDA+LDA, and BTTDA+LDA.

@@

Nevertheless, when considering the relative improvement over the original
HODA model, our results show that BTTDA+LDA performs consistently on par or
higher than HODA+LDA, with performances of respectively $86.43>83.25$ and
$85.08>82.57$ for \ac{erp} datasets and $55.13>53.51$ and $80.60<80.88$ for
\ac{mi} datasets.
The classification score for BTTDA+LDA was also slightly
higher than that of
restricted PARAFACDA+LDA for the ERP datasets ($86.43>86.19$ and $85.06>84.85$
respectively) and one of the MI datasets ($55.13>54.34$ and $80.60<80.88$
respectively), but these results were not significant, and further studies with
more datasets and :subjects should show whether this holds.
\Cref{fig:blocks} shows that there is added value in finding extra blocks
over the first HODA block.
While no proof is given, we notice that NMSE monotonically decreases, indicating that
since the training score keeps increasing, this suggests that
eventually all the variation in the signal will be explained by the model
while still extracting features that are maximally discriminant.
Eventually, the number of blocks will reach a point of diminishing validation
score returns, when adding extra features to the decision classifier increases
its risk of overfitting instead of adding extra useful discriminatory
information.
Together with the improved classification scores presented, these results
point to the potential of our more flexible model block-term or
its special PARAFAC-structured case over a Tucker-structured model given proper block and
rank selection.
Since the optimal ranks for HODA+LDA were also determined through
cross-validation, BTTDA+LDA can improve over the first HODA block.
If validation shows that this is not possible, such as is the case for
BNCI2014-004, the BTTDA model is truncated to the first HODA block and little
performance is lost.

@@

Via an accompanying heuristic model selection procedure, \ac{bci} decoders using
\ac{bttda} feature extraction can significantly outperform decoders based on
\ac{hoda} and reach state-of-the-art decoding performance on event-related potential
problems (second-order tensors) and scores on par with or higher than \ac{hoda} in motor
imagery problems (third-order tensors).
Moving from the rigid Tucker tensor structure of \ac{hoda} to the more flexible
block-term structure shifts the problem from finding optimally constrained
multilinear projections to model and feature selection.
